### 2025-05-29 Manuscript/Lower bounds of the success probability in quantum state exclusion for general ensembles.tex
Stuff

@@

In this section we introduce some mathematical formalism of the postulates of Quantum Mechanics as well as some relevant objects in order to introduce the mathematical framework of \gls{qse} in a \gls{sdp} form. Additionally we define the Haar measure and discuss its importance for a proper generation of generic ensembles.

@@

This project is entirely situated within the quantum realm. Therefore, we begin by contextualizing the foundational postulates, notation, and elements upon which the rest of the project relies on.

@@

Let $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ be an ensemble of $n$ quantum states, where each $\rho_i$ denotes a pure state density matrix\footnote{This formulation hols true for mixed states but the project will only discuss the pure state scenario.}, i.e., $\rho_i = \ket{\psi_i}\bra{\psi_i}$, and $\eta_i$ represents the prior probability of occurrence of the state $\rho_i$. Let $\tilde{\rho}$ be a state from this ensemble. Our objective is to develop a procedure to identify another state $\rho_j \in \left\{\rho_i\right\}_{i=1}^n$, such that $\rho_j \neq \tilde{\rho}$.

@@

To ensure completeness, we introduce an additional \gls{povm} element $\Pi_?$ representing an inconclusive result,
\begin{align*}
\Pi_? = \mathds{1} - \sum_{i=1}^n \Pi_i.
\end{align*}
If the measurement yields the outcome $\Pi_?$, i.e., the "?" label, the result is inconclusive.

@@

\subsection{Gram Matrix Formulation}\label{sectionGramMatrixFormulation}
The Gram matrix formulation recasts the \gls{sdp} problems associated with the exclusion task in a more basis-independent manner. This approach reveals that the exclusion tasks for certain ensembles can be equivalent to those of others. As such, we adopt this formulation in our analysis, as it offers valuable insight into the underlying relationships between the quantum states in the ensemble.

Given an ensemble $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ of $n$ quantum states we define the $\mathcal{G} \in \mathbb{C}^{n \times n}$ be the \emph{Gram matrix} of the ensemble, as the $n \times n$ positive semidefinite Hermitian matrix given by,
\begin{equation*}
	\mathcal{G} =
	\begin{pmatrix}
		\eta_1 & \sqrt{\eta_1\eta_2}\braket{\psi_1|\psi_2} & \dots & \sqrt{\eta_1\eta_n}\braket{\psi_1|\psi_n} \\
		\sqrt{\eta_2\eta_1}\braket{\psi_2|\psi_1} & \eta_2 & \dots & \sqrt{\eta_2\eta_n}\braket{\psi_2|\psi_n} \\
		\vdots & \vdots & \ddots & \vdots \\
		\sqrt{\eta_n\eta_1}\braket{\psi_n|\psi_1} & \sqrt{\eta_n\eta_2}\braket{\psi_n|\psi_2} & \dots & \eta_n
	\end{pmatrix},
\end{equation*}
that is, $\mathcal{G}_{i,j} = \sqrt{\eta_i\eta_j}\braket{\psi_i|\psi_j}$. Since all states are normalized, the diagonal elements are equal to $\eta_i$.

@@

As we have mentioned before, analytical results have been found for the exclusion task in both protocols \gls{me} and \gls{ze} for ensembles that exhibit a certain degree of symmetry\cite{MainPaper}, that is for \emph{group-generated} ensembles. In this section we present a proper definition for a group-generated ensemble as well as a brief discussion on this ensembles generated by a cyclic group, a key type of grop-generated ensembles in our analysis.

@@

For instance, let $\mathcal{U}$ be a unitary operator such that $\mathcal{U}^n = \mathds{1}$ (i.e., $\mathcal{U}$ generates a cyclic group of order $n$). Then, the set of states
\begin{align*}
	\left\{ \mathcal{U}^i\ket{\psi} \right\}_{i=0}^{n-1},
\end{align*}
forms a group generated ensemble based on the cyclic group $\mathbb{Z}/n\mathbb{Z}$. This type of ensemble is of particular interest in our study and will be explored in more detail at the end of this section. Additionally, in Appendix \ref{appendixComputationGroupGeneratedGramMatrices} we present a general procedure to compute the Gram Matrix for any group-generated ensemble.

@@

Here we discuss the particular case for a $\mathbb{Z}/n\mathbb{Z}$ group-generated ensemble a key type of ensemble for our project.\footnote{The relevance of this type of ensembles relies on the fact that for a prime number of states, the possible group-generated group ensemble } Let $\mathcal{U} \in U(n)$ be an $n \times n$ unitary matrix satisfying $\mathcal{U}^n = \mathds{1}$, and let $\ket{\psi}$ be the seed state. The Gram matrix $\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}$ elements associated with the ensemble $\left\{\mathcal{U}^i\ket{\psi}\right\}_{i=0}^{n-1}$ are,
\begin{align*}
	\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}_{i,j} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^i\psi | \mathcal{U}^j\psi} = \sqrt{\eta_i \eta_j}\braket{\psi | \mathcal{U}^{j-i} | \psi} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^{j-i}}_{\psi},
\end{align*}

@@

Note that the Gram matrix is Hermitian, as expected, since
\begin{align*}
    \braket{\mathcal{U}^{-j}}_{\psi} = \braket{\psi | \mathcal{U}^{-j} | \psi} = \left(\braket{\psi | \mathcal{U}^j | \psi}\right)^* = \braket{\mathcal{U}^j}_{\psi}^*.
\end{align*}
Additionally, by using the identity $\mathcal{U}^n = \mathds{1}$, we can simplify expressions such as,
\begin{align*}
	\braket{\mathcal{U}^{-n+i}}_\psi = \braket{\psi | \mathcal{U}^{-n+i} | \psi} = \braket{\psi | \mathcal{U}^i | \psi} = \braket{\mathcal{U}^i}_\psi.
\end{align*}
Thus, we may write,
\begin{align*}
	\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}} = \begin{pmatrix}
 \eta_1 & \sqrt{\eta_1 \eta_2}\braket{\mathcal{U}}_{\psi} & \sqrt{\eta_1 \eta_3}\braket{\mathcal{U}^2}_{\psi} & \cdots & \sqrt{\eta_1 \eta_n}\braket{\mathcal{U}}^*_{\psi} \\
 \sqrt{\eta_2 \eta_1}\braket{\mathcal{U}}^*_{\psi} & \eta_2 & \sqrt{\eta_2 \eta_3}\braket{\mathcal{U}}_{\psi} & \cdots & \sqrt{\eta_2 \eta_n}\braket{\mathcal{U}}^*_{\psi} \\
 \sqrt{\eta_1 \eta_3}\braket{\mathcal{U}^2}^*_{\psi} & \sqrt{\eta_2 \eta_3}\braket{\mathcal{U}}^*_{\psi} & \eta_3 & \cdots & \sqrt{\eta_3 \eta_n}\braket{\mathcal{U}}^*_{\psi} \\
 \vdots & \vdots & \vdots & \ddots & \vdots \\
 \sqrt{\eta_1 \eta_n}\braket{\mathcal{U}}_{\psi} & \sqrt{\eta_2 \eta_n}\braket{\mathcal{U}^{2}}_{\psi} & \sqrt{\eta_3 \eta_n}\braket{\mathcal{U}^{3}}_{\psi} & \cdots & \eta_n
\end{pmatrix}.
\end{align*}

@@

The Fourier basis for an $n \times n$ circulant matrix consists of the set of vectors $\{\omega_i\}_{i=1}^n$, where
\begin{align*}
	\omega_i = \frac{1}{\sqrt{n}} (1, \gamma^i, \gamma^{2i}, \ldots, \gamma^{i(n-1)})^T,
\end{align*}
and
\begin{align*}
	\gamma = e^{\frac{2\pi i}{n}}.
\end{align*}

@@

Significant progress has been made in the study of the state exclusion task for group-generated ensembles. In particular, the work \emph{Quantum State Exclusion for Group-Generated Ensembles of Pure States}\cite{MainPaper} presents exact analytical expressions for the success and error probabilities under both the Minimum Error (\gls{me}) and Zero Error (\gls{ze}) exclusion protocols. In this section we present this results and comment some key aspects of this results regarding our project, in particular we present the concept of \emph{perfect exclusion region} and we discuss the case of ensembles with a prime number of states.

Let $\mathcal{G}^G$ denote the Gram matrix of a group-generated ensemble $\left\{ \left( \rho_i, \frac{1}{|G|} \right) \right\}_{i=1}^{|G|}$, where $G$ is a finite group of size $|G|$, and let $\{\lambda_i\}_{i=1}^{|G|}$ be the set of eigenvalues of $\mathcal{G}^G$, the Gram Matrix of the group. Assume the eigenvalues $\left\{\frac{\lambda_i}{|G|}|1\leq i \leq n\right\}$ are ordered increasingly,
\begin{align*}
	i \leq j \quad \Leftrightarrow \quad \lambda_i \leq \lambda_j, \quad \forall i,j \in \{1, \dots, |G|\},
\end{align*}
i.e., $\frac{\lambda_1}{|G|}$ is the smallest and $\lambda_{|G|}$ the largest eigenvalue. Notice that since we are in the equal probability scenario, we can factor out a $\frac{1}{|G|}$ factor from the Gram Matrix and consider the matrix $|G|\mathcal{G}^G$ instead of $\mathcal{G}^G$, i.e. the eigenvalues of this matrix simply are $\left\{\lambda_i\right\}_{i=1}^n$. This is a standard choice for simplicity, when we dealing with equal prior probabilities.

@@

Furthermore, we might highlight a key aspect of the analytical solutions for the \gls{qse} task form \gls{me} and \gls{ze} protocols. These results depend solely on the eigenvalues of the Gram matrix and are independent of the specific group structure generating the ensemble. Hence, let the Gram matrix of an ensemble decompose as,
\begin{align*}
	\mathcal{G} = U D U^\dagger,
\end{align*}
where $U$ is a unitary matrix and $D$ is the diagonal matrix of eigenvalues. We then encounter two possibilities,
\begin{itemize}
	\item The eigenvalues satisfy Eq.~\eqref{equationPerfectExclusionCondition}, placing the ensemble in the \emph{perfect exclusion} regime.
	\item The eigenvalues do not satisfy Eq.~\eqref{equationPerfectExclusionCondition}, placing the ensemble in the \emph{non-perfect exclusion} regime.
\end{itemize}

@@

All the code used in this project is hosted in the GitHub repository \cite{GitHub}, where additional documentation is available detailing the functionality and structure of each script and class involved in the implementation. The numerical analysis is performed by generating ensembles uniformly distributed taking into account all possibilities and then solving the \gls{sdp} problems presented in Section \ref{sectionFormulationOfTheProblem}.

@@

A Numpy \gls{rng}\cite{numpyRNG} produces a set of $n$ non-negative real numbers $\{\lambda_i\}_{i=1}^n$ that are normalized to satisfy the trace condition,
\begin{align*}
	\sum_{i=1}^n \lambda_i = 1.
\end{align*}
These values are used as the eigenvalues of the Gram matrix, forming the diagonal matrix $D = \operatorname{diag}(\lambda_1, \dots, \lambda_n)$.

@@

The begining of the analysis consists in showing up the key differences between \gls{me} and \gls{ze} protocols, and the concordance of the \gls{sdp}'s numerical results with the analytical results given in \ref{sectionPreviousResults}.

@@

Here we present a comparison between both protocols for \gls{ze} and \gls{me} giving insights on the protocols's essence and contextualising the problem in a visual way.

In Figure (\ref{FigureQSEMEZ3ZHeatmap})  we presents the \gls{me} (left) and \gls{ze} (right) protocols results for the \gls{qse} task. In particular Figure (\ref{FigureQSEMEZ3ZHeatmap}) shows two heatmaps of the average successful exclusion probability under the Minimum Error (\gls{me}) protocol for a $\mathbb{Z}/3\mathbb{Z}$ group-generated ensemble. Probabilities are shown as functions of the modulus $|c|$ ($x$-axis) and phase $\theta$ ($y$-axis) of the overlap $c=|c|^{i\theta}$, which fully determine the Gram matrix. Notice we take equal prior probabilities, as discussed in the analytical results in Section (\ref{sectionPreviousResults}).

@@

In this section, we present the results for ensembles generated by the group $\mathbb{Z}/3\mathbb{Z}$, analyzed across the spectrum of eigenvalues, i.e. the analogus results from the previos section in the eigenvalues spectum. We also explicitly show the conditions for perfect exclusion, we give a proper analysis of this regime, and provide an initial exploration of the results for general ensembles being low bounded by the group-generated results of three quantum states.

@@

Naturally we can prove the perfect exclusion zone to be a subset of valid Gram matrixes i.e. $\Delta\subset \Omega$. Then, the non-perfect exclusion regime is the remaining part of the triangle i.e.,
\begin{align*}
	\text{Non-perfect exclusion regime}=\Omega\setminus \Delta.
\end{align*}
The limit case scenarios corresponding to $\lambda_i=3$ and $\lambda_j,\lambda_k=0$ are located in the vertex of the triangle\footnote{Remember this case correspond to sets with the same state repeated 3 times discussed in Section (\ref{sectionLimitCaseScenarios})}. Additionally, the cases corresponding to the edges are those with one state repeated 2 times and one different from the other 2. This ensembles can also be underestood as a limit case scenario systems.

@@

\subsection{Distribution of the general ensembles success exclusion probability}
Here we present the main results of the project. Hence we show an analytical results for generic ensembles and how the success exclusion probability ensembles are distributed over the results for group-generate ensembles, i.e. the lower bound.

We know every Gram Matrix $\mathcal{G}$ can be written as $\mathcal{G}=UDU^\dagger$. By fixing the eigenvalues $\{g_i\}_{i=1}^n$, i.e. the elements of the diagonal matrix $D$\footnote{Note we have changed the notation respect the previous  eigenvalues since before we were taking the eigenvalues of the Gram Matrix where factoring out $\frac{1}{n}$ in the equal prior probabilities scenario. This eigenvalues were normalised to $n$, so we change the notation for avoiding any possible confusion.}, as normalised elements,
\begin{align*}
\sum_{i=1}^ng_i=1
\end{align*}

With this elements we can compute the lower bound given in section (\ref{sectionPreviousResults}) and find whether they are in the perfect exclusion region or not. If they are, any Gram matrix generated via $G=UDU^\dagger$ would represent a set of ensembles such that perfect exclusion can be archived and the lower bound would be the exact result. Otherwise, if we are in the non-perfect exclusion region by interchanging the unitary matrices $U$ we can generate different sets of Gram matrices such that

Since the algorithm introduced by Francesco Mezzadri in \cite{UnitaryMatricesGeneration}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth, trim={1.0cm 0.3cm 2.4cm 1.5cm}, clip]{../Plots/ExclusionMinimumErrorRandomDistributionZ3Prob0.962.pdf}
	\caption{Hisotgram depicting the distribution of the generated Gram Matrices over the }
	\label{FigureDistZ5ME0.962}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth, trim={1.0cm 0.3cm 2.4cm 1.5cm}, clip]{../Plots/ExclusionZeroErrorRandomDistributionZ3Prob0.5.pdf}
	\caption{Hisotgram of the group genearation }
	\label{FigureDistZ5ZE0.5}
\end{figure}

@@

It is often useful to consider the \emph{dual} version of a \gls{sdp} problem, as it provides both conceptual insights and practical advantages for verifying optimality.

>>>>

\interfootnotelinepenalty=10000

@@

\hspace{20pt}In this section we introduce some mathematical formalism of the postulates of Quantum Mechanics as well as some relevant objects in order to introduce the mathematical framework of \gls{qse} in a \gls{sdp} form. Additionally we define the Haar measure and discuss its importance for a proper generation of generic ensembles.

@@

\hspace{20pt}This project is entirely situated within the quantum realm. Therefore, we begin by contextualizing the foundational postulates, notation, and elements upon which the rest of the project relies on.

@@

\hspace{20pt}Let $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ be an ensemble of $n$ quantum states, where each $\rho_i$ denotes a pure state density matrix\footnote{This formulation hols true for mixed states but the project will only discuss the pure state scenario.}, i.e., $\rho_i = \ket{\psi_i}\bra{\psi_i}$, and $\eta_i$ represents the prior probability of occurrence of the state $\rho_i$. Let $\tilde{\rho}$ be a state from this ensemble. Our objective is to develop a procedure to identify another state $\rho_j \in \left\{\rho_i\right\}_{i=1}^n$, such that $\rho_j \neq \tilde{\rho}$.

@@

To ensure completeness, we introduce an additional \gls{povm} element $\Pi_?$ representing an inconclusive result,
\begin{align*}
\Pi_? = \mathds{1} - \sum_{i=1}^n \Pi_i.
\end{align*}

If the measurement yields the outcome $\Pi_?$, i.e., the "?" label, the result is inconclusive.

@@

\subsection{Gram Matrix Formulation}\label{sectionGramMatrixFormulation}
\hspace{20pt}The Gram matrix formulation recasts the \gls{sdp} problems associated with the exclusion task in a more basis-independent manner. This approach reveals that the exclusion tasks for certain ensembles can be equivalent to those of others. As such, we adopt this formulation in our analysis, as it offers valuable insight into the underlying relationships between the quantum states in the ensemble.

Given an ensemble $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ of $n$ quantum states we define the $\mathcal{G} \in \mathbb{C}^{n \times n}$ be the \emph{Gram matrix} of the ensemble, as the $n \times n$ positive semidefinite Hermitian matrix given by,
\begin{equation*}
	\mathcal{G} =
	\begin{pmatrix}
		\eta_1 & \sqrt{\eta_1\eta_2}\braket{\psi_1|\psi_2} & \dots & \sqrt{\eta_1\eta_n}\braket{\psi_1|\psi_n} \\
		\sqrt{\eta_2\eta_1}\braket{\psi_2|\psi_1} & \eta_2 & \dots & \sqrt{\eta_2\eta_n}\braket{\psi_2|\psi_n} \\
		\vdots & \vdots & \ddots & \vdots \\
		\sqrt{\eta_n\eta_1}\braket{\psi_n|\psi_1} & \sqrt{\eta_n\eta_2}\braket{\psi_n|\psi_2} & \dots & \eta_n
	\end{pmatrix},
\end{equation*}

that is, $\mathcal{G}_{i,j} = \sqrt{\eta_i\eta_j}\braket{\psi_i|\psi_j}$. Since all states are normalized, the diagonal elements are equal to $\eta_i$.

@@

\hspace{20pt}As we have mentioned before, analytical results have been found for the exclusion task in both protocols \gls{me} and \gls{ze} for ensembles that exhibit a certain degree of symmetry\cite{MainPaper}, that is for \emph{group-generated} ensembles. In this section we present a proper definition for a group-generated ensemble as well as a brief discussion on this ensembles generated by a cyclic group, a key type of grop-generated ensembles in our analysis.

@@

For instance, let $\mathcal{U}$ be a unitary operator such that $\mathcal{U}^n = \mathds{1}$ (i.e., $\mathcal{U}$ generates a cyclic group of order $n$). Then, the set of states
\begin{align*}
	\left\{ \mathcal{U}^i\ket{\psi} \right\}_{i=0}^{n-1},
\end{align*}

forms a group generated ensemble based on the cyclic group $\mathbb{Z}/n\mathbb{Z}$. This type of ensemble is of particular interest in our study and will be explored in more detail at the end of this section. Additionally, in Appendix \ref{appendixComputationGroupGeneratedGramMatrices} we present a general procedure to compute the Gram Matrix for any group-generated ensemble.

@@

\hspace{20pt}Here we discuss the particular case for a $\mathbb{Z}/n\mathbb{Z}$ group-generated ensemble a key type of ensemble for our project.\footnote{The relevance of this type of ensembles relies on the fact that for a prime number of states, the possible group-generated group ensemble } Let $\mathcal{U} \in U(n)$ be an $n \times n$ unitary matrix satisfying $\mathcal{U}^n = \mathds{1}$, and let $\ket{\psi}$ be the seed state. The Gram matrix $\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}$ elements associated with the ensemble $\left\{\mathcal{U}^i\ket{\psi}\right\}_{i=0}^{n-1}$ are,
\begin{align*}
	\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}_{i,j} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^i\psi | \mathcal{U}^j\psi} = \sqrt{\eta_i \eta_j}\braket{\psi | \mathcal{U}^{j-i} | \psi} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^{j-i}}_{\psi},
\end{align*}

@@

Note that the Gram matrix is Hermitian, as expected, since
\begin{align*}
    \braket{\mathcal{U}^{-j}}_{\psi} = \braket{\psi | \mathcal{U}^{-j} | \psi} = \left(\braket{\psi | \mathcal{U}^j | \psi}\right)^* = \braket{\mathcal{U}^j}_{\psi}^*.
\end{align*}

Additionally, by using the identity $\mathcal{U}^n = \mathds{1}$, we can simplify expressions such as,
\begin{align*}
	\braket{\mathcal{U}^{-n+i}}_\psi = \braket{\psi | \mathcal{U}^{-n+i} | \psi} = \braket{\psi | \mathcal{U}^i | \psi} = \braket{\mathcal{U}^i}_\psi.
\end{align*}

Thus, we may write,
\begin{align*}
	\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}} = \begin{pmatrix}
 \eta_1 & \sqrt{\eta_1 \eta_2}\braket{\mathcal{U}}_{\psi} & \sqrt{\eta_1 \eta_3}\braket{\mathcal{U}^2}_{\psi} & \cdots & \sqrt{\eta_1 \eta_n}\braket{\mathcal{U}}^*_{\psi} \\
 \sqrt{\eta_2 \eta_1}\braket{\mathcal{U}}^*_{\psi} & \eta_2 & \sqrt{\eta_2 \eta_3}\braket{\mathcal{U}}_{\psi} & \cdots & \sqrt{\eta_2 \eta_n}\braket{\mathcal{U}}^*_{\psi} \\
 \sqrt{\eta_1 \eta_3}\braket{\mathcal{U}^2}^*_{\psi} & \sqrt{\eta_2 \eta_3}\braket{\mathcal{U}}^*_{\psi} & \eta_3 & \cdots & \sqrt{\eta_3 \eta_n}\braket{\mathcal{U}}^*_{\psi} \\
 \vdots & \vdots & \vdots & \ddots & \vdots \\
 \sqrt{\eta_1 \eta_n}\braket{\mathcal{U}}_{\psi} & \sqrt{\eta_2 \eta_n}\braket{\mathcal{U}^{2}}_{\psi} & \sqrt{\eta_3 \eta_n}\braket{\mathcal{U}^{3}}_{\psi} & \cdots & \eta_n
\end{pmatrix}.
\end{align*}

@@

The Fourier basis for an $n \times n$ circulant matrix consists of the set of vectors $\{\omega_i\}_{i=1}^n$, where
\begin{align*}
	\omega_i = \frac{1}{\sqrt{n}} (1, \gamma^i, \gamma^{2i}, \ldots, \gamma^{i(n-1)})^T,
\end{align*}

and
\begin{align*}
	\gamma = e^{\frac{2\pi i}{n}}.
\end{align*}

@@

\hspace{20pt}Significant progress has been made in the study of the state exclusion task for group-generated ensembles. In particular, the work \emph{Quantum State Exclusion for Group-Generated Ensembles of Pure States}\cite{MainPaper} presents exact analytical expressions for the success and error probabilities under both the Minimum Error (\gls{me}) and Zero Error (\gls{ze}) exclusion protocols. In this section we present this results and comment some key aspects of this results regarding our project, in particular we present the concept of \emph{perfect exclusion region} and we discuss the case of ensembles with a prime number of states.

Let $\mathcal{G}^G$ denote the Gram matrix of a group-generated ensemble $\left\{ \left( \rho_i, \frac{1}{|G|} \right) \right\}_{i=1}^{|G|}$, where $G$ is a finite group of size $|G|$, and let $\{\lambda_i\}_{i=1}^{|G|}$ be the set of eigenvalues of $\mathcal{G}^G$, the Gram Matrix of the group. Assume the eigenvalues $\left\{\frac{\lambda_i}{|G|}|1\leq i \leq n\right\}$ are ordered increasingly,
\begin{align*}
	i \leq j \quad \Leftrightarrow \quad \lambda_i \leq \lambda_j, \quad \forall i,j \in \{1, \dots, |G|\},
\end{align*}

i.e., $\frac{\lambda_1}{|G|}$ is the smallest and $\lambda_{|G|}$ the largest eigenvalue. Notice that since we are in the equal probability scenario, we can factor out a $\frac{1}{|G|}$ factor from the Gram Matrix and consider the matrix $|G|\mathcal{G}^G$ instead of $\mathcal{G}^G$, i.e. the eigenvalues of this matrix simply are $\left\{\lambda_i\right\}_{i=1}^n$. This is a standard choice for simplicity, when we dealing with equal prior probabilities.

@@

Furthermore, we might highlight a key aspect of the analytical solutions for the \gls{qse} task form \gls{me} and \gls{ze} protocols. These results depend solely on the eigenvalues of the Gram matrix and are independent of the specific group structure generating the ensemble. Hence, let the Gram matrix of an ensemble decompose as,
\begin{align*}
	\mathcal{G} = U D U^\dagger,
\end{align*}

where $U$ is a unitary matrix and $D$ is the diagonal matrix of eigenvalues. We then encounter two possibilities,
\begin{itemize}
	\item The eigenvalues satisfy Eq.~\eqref{equationPerfectExclusionCondition}, placing the ensemble in the \emph{perfect exclusion} regime.
	\item The eigenvalues do not satisfy Eq.~\eqref{equationPerfectExclusionCondition}, placing the ensemble in the \emph{non-perfect exclusion} regime.
\end{itemize}

@@

\hspace{20pt}All the code used in this project is hosted in the GitHub repository \cite{GitHub}, where additional documentation is available detailing the functionality and structure of each script and class involved in the implementation. The numerical analysis is performed by generating ensembles uniformly distributed taking into account all possibilities and then solving the \gls{sdp} problems presented in Section \ref{sectionFormulationOfTheProblem}.

@@

A Numpy \gls{rng}\cite{numpyRNG} produces a set of $n$ non-negative real numbers $\{\lambda_i\}_{i=1}^n$ that are normalized to satisfy the trace condition,
\begin{align*}
	\sum_{i=1}^n \lambda_i = 1.
\end{align*}

These values are used as the eigenvalues of the Gram matrix, forming the diagonal matrix $D = \operatorname{diag}(\lambda_1, \dots, \lambda_n)$.

@@

\hspace{20pt}The begining of the analysis consists in showing up the key differences between \gls{me} and \gls{ze} protocols, and the concordance of the \gls{sdp}'s numerical results with the analytical results given in \ref{sectionPreviousResults}.

@@

\hspace{20pt}Here we present a comparison between both protocols for \gls{ze} and \gls{me} giving insights on the protocols's essence and contextualising the problem in a visual way.

In Figure (\ref{FigureQSEMEZ3ZHeatmap})  we presents the \gls{me} (left) and \gls{ze} (right) protocols results for the \gls{qse} task. In particular Figure (\ref{FigureQSEMEZ3ZHeatmap}) shows two heatmaps of the average successful exclusion probability under the Minimum Error (\gls{me}) protocol for a $\mathbb{Z}/3\mathbb{Z}$ group-generated ensemble. Probabilities are shown as functions of the modulus $|c|$ ($x$-axis) and phase $\theta$ ($y$-axis) of the overlap $c=|c|^{i\theta}$, which fully determines the Gram matrix. Notice we take equal prior probabilities, as discussed in the analytical results in Section (\ref{sectionPreviousResults}).

@@

\hspace{20pt}In this section, we present the results for ensembles generated by the group $\mathbb{Z}/3\mathbb{Z}$, analyzed across the spectrum of eigenvalues, i.e. the analogus results from the previos section in the eigenvalues spectum. We also explicitly show the conditions for perfect exclusion, we give a proper analysis of this regime, and provide an initial exploration of the results for general ensembles being low bounded by the group-generated results of three quantum states.

@@

Naturally we can prove the perfect exclusion zone to be a subset of valid Gram matrixes i.e. $\Delta\subset \Omega$. Then, the non-perfect exclusion regime is the remaining part of the triangle i.e.,
\begin{align*}
	\text{Non-perfect exclusion regime}=\Omega\setminus \Delta.
\end{align*}

The limit case scenarios corresponding to $\lambda_i=3$ and $\lambda_j,\lambda_k=0$ are located in the vertex of the triangle\footnote{Remember this case correspond to sets with the same state repeated 3 times discussed in Section (\ref{sectionLimitCaseScenarios})}. Additionally, the cases corresponding to the edges are those with one state repeated 2 times and one different from the other 2. This ensembles can also be underestood as a limit case scenario systems.

@@

\subsection{Distribution of the general ensembles success exclusion probability}

\hspace{20pt}In this section, we present the main results of the project. Specifically, we provide analytical results for generic ensembles and examine how the success probabilities for quantum state exclusion are distributed in relation to the lower bound defined by group-generated ensembles.

We begin by recalling that any Gram matrix $ \mathcal{G} $ can be decomposed as $ \mathcal{G} = U D U^\dagger $, where $ D $ is a diagonal matrix containing the eigenvalues $\{g_i\geq0\}_{i=1}^n $. We fix these eigenvalues such that
\begin{align*}
\sum_{i=1}^n g_i = 1,
\end{align*}
ensuring they are normalized.\footnote{Note that this eigenvalue notation differs from that used earlier in the document, where the eigenvalues of the Gram matrix were considered with an overall normalization factor of \( \frac{1}{n} \), assuming equal prior probabilities and were normalised to $n$. To avoid confusion, we change the notation for normalised eigenvalues that sum up to 1 for the general prior probabilities case.}

With this eigenvalue spectrum, we compute the analytical lower bound introduced in Section~\ref{sectionPreviousResults} and determine whether it falls within the perfect exclusion region. If it does, then any Gram matrix generated as $ \mathcal{G} = U D U^\dagger $ represents an ensemble for which perfect exclusion is feasible, and the lower bound is in fact the success probability, i.e. both are 1. Conversely, if the lower bound lies outside the perfect exclusion region, we explore different unitary matrices $ U $ to generate alternative Gram matrices that share the same eigenvalues, thus preserving the lower bound, while corresponding to distinct ensembles.

We use the algorithm proposed in~\cite{UnitaryMatricesGeneration} to generate unitary matrices sampled uniformly from \( U(n) \). This ensures a comprehensive exploration of the space of possible Gram matrices, i.e., ensembles, corresponding to a fixed lower bound.

It is important to highlight that this method also implicitly generates random prior probabilities for each state, since the diagonal elements of the Gram matrix are the priors probabilities, and generally will differ from one to another. Nevertheless, the analytical lower bound remains valid, as we will show in the following results.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth, trim={1.0cm 0.3cm 2.4cm 1.5cm}, clip]{../Plots/ExclusionMinimumErrorRandomDistributionZ3Prob0.962.pdf}
	\caption{Hisotgram depicting the distribution of generated Gram Matrices via the success probability for \gls{me} protocol with a lower bound of 0.962 for 5 states ensembles of random priors probabilities.}
	\label{FigureDistZ5ME0.962}
\end{figure}

Figure (\ref{FigureDistZ5ME0.962}) shows the distribution of success probabilities for quantum state exclusion using the \gls{me} protocol computed via semidefinite programming (\gls{sdp}) for 5 states ensembles and random priors probabilities. The histogram's $x$-axis represents the exclusion success probability, while the $y$-axis indicates the frequency of occurrence across sampled ensembles. Additionally the dashed blue line marks the analytical lower bound.

We emphasize that each computed result has been verified not to fall below this analytical lower bound. Although a few discrepancies were observed, these occurred only at the fifth decimal place, where numerical precision limitations in the \gls{sdp} solver can affect reliability. These cases are treated as inconclusive due to numerical problems, and no inconsistencies have been found exceeding a deviation in the fourth decimal.

From the results in Figure (\ref{FigureDistZ5ME0.962}), we observe the ensembles are broadly distributed across a range of success probabilities. However, a significant clustering occurs near a success probability of 1 under the \gls{me} protocol, indicating that perfect exclusion is achievable for a substantial subset of these randomly generated ensembles.

These ensembles were generated uniformly by sampling unitary matrices $ U $ from $U(n)$, which determine the Gram matrices via the decomposition $\mathcal{G} = U D U^\dagger$. Ensembles with Gram matrices structurally similar to those of group-generated ensembles, particularly those diagonalized in the Fourier basis\footnote{Recall that we are working with ensembles of a prime number of states, for which the only known analytical result corresponds to the Gram matrix generated by $\mathbb{Z}/n\mathbb{Z}$, which is diagonalized in the Fourier basis. A more complete study would consider other group-generated cases.}, yield lower success probabilities.

Interestingly, there exists a threshold of deviation from the Fourier basis beyond which all ensembles yield a success probability of 1. In other words, beyond a certain \emph{distance} from the group-generated case, i.e. the Fourier basis, the ensembles become sufficiently distinguishable to allow perfect exclusion, regardless of the specific basis in which their Gram matrices are diagonalised.

At this stage we can observe the crucial importance of the generation of the unitary matrices uniformly covering  $U(n)$, i.e. being generated according to the Haar measure. If we could not ensure we are exploring all the possible ensembles this observations and the study of the generic ensembles case would be completely pointless since we can overrepresent one specific kind of ensambles respect to others.

In Figure (\ref{FigureDistZ5ZE0.5}) we present the analogous results for the \gls{ze} protocol for ensembles consisting in 5 states with eigenvalues that lead to an analytical lower bound for success probability of $0.5$ in this protocol.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth, trim={1.0cm 0.3cm 2.4cm 1.5cm}, clip]{../Plots/ExclusionZeroErrorRandomDistributionZ3Prob0.5.pdf}
	\caption{Hisotgram depicting the distribution of generated Gram Matrices via the success probability for \gls{ze} protocol with a lower bound of 0.5 for 5 states ensembles of random priors probabilities.}
	\label{FigureDistZ5ZE0.5}
\end{figure}

Thus, we observe a similar behaviour respect to Figure (\ref{FigureDistZ5ME0.962}), still we attribute the pick of ensembles where perfect exclusion is feasible to the threshold of the \emph{distance} from the group generated case as previously discussed. Still some remarkable aspect of this protocol is the constant is the constant decay in the frequency for higher exclusion probabilities.

As before, no ensembles yields lower success exclusion probabilities than the lower bound, proving this as a lower bound of the generic case scenario.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.49\textwidth, trim={1.0cm 0.3cm 2.4cm 1.5cm}, clip]{../Plots/ExclusionMinimumErrorRandomDistributionZ7Prob0.924.pdf}
	\includegraphics[width=0.49\textwidth, trim={1.0cm 0.3cm 2.4cm 1.5cm}, clip]{../Plots/ExclusionMinimumErrorRandomDistributionZ11Prob0.926.pdf}
	\caption{Hisotgram depicting the distribution of generated Gram Matrices via the success probability for \gls{me} protocol with a lower bound of 0.92 for 7 states (left) and 11 states (right) ensembles of random priors probabilities.}
	\label{FigureDistZ5ZE0.5}
\end{figure}

@@

\hspace{20pt}It is often useful to consider the \emph{dual} version of a \gls{sdp} problem, as it provides both conceptual insights and practical advantages for verifying optimality.
